Think deep for reef resilience

### 2025-08-12 paper/paper_draft.qmd
Expand limitation section
I think the connectivity angle warrants further description here too @eco-ben

Additionally, it is currently not possible to obtain highly accurate and precise estimates
of reef depth at the GBR-scale. Data adopted for the given assessment are said to be
accurate up to 15m from Mean Sea Level [@GBRMPA2020], however there are many sources of
uncertainty in the depth estimates used for this study. Satellite-derived estimates are
subject to different bottom types (rock, sand, seagrass, etc) and how these affect
reflectance, the water penetration depth of satellite sensors, influence of local conditions
(e.g., turbidity and wave activity), among many other factors.

In addition to the usual challenges of deriving data from satellite imagery, it is
acknowledged that there is high potential for misalignment between the bathymetry data and
GBRMPA reef outlines. Hammerton and Lawrey [-@hammerton], for example, identified 120 false
positives, where features identified as reefs were in actuality sand banks or rocky reefs
(rather than coral reefs). Reefs with bathymetry data of less than 5% of its indicative area
were excluded from analysis to ensure analyses were not conducted on reefs with no data
coverage (see @sfig-depth-cutoff). The existing GBRMPA reef outlines were adopted to align
with other ongoing modelling and analyses conducted as part of the Reef Restoration and
Adaptation Program \[RRAP; <https://gbrrestoration.org>\] **\[TODO: PROPER CITATION OF
RRAP\]**. The median depths for each indicative reef area were extracted from bathymetry
data using zonal statistics and depth values beyond the 15m range were included. The
resulting depth estimates are therefore highly uncertain and are only adopted as a useful
indication for the purpose of the presented analysis. They are not intended to be an
accurate and precise report of reef depths across the GBR.

**\[TODO - rewrite/expand paragraphs below for flow\]** As ADRIA-CoralBlox has not been
validated at the GBR-wide scale, we aim to assess the relative influence of reef
characteristics on resilience under future change rather than provide spatially explicit
predictions of reef outcomes into the future. Using this approach we suggest that median
reef depth is an important factor for reef resilience under thermal stress. Improvements in
remote sensing approaches for broad scale bathymetry measurement, and understanding of how
depth impacts experienced thermal stress on reefs will further enhance investigations of
depth as a critical factor for reef resilience. Additional analyses could build on the
findings of this study by assessing the impacts of prioritising reef depth in reef
restoration decision making. The ADRIA platform is developed as a modelling tool to
incorporate multi-criteria decision analysis (MCDA) and ecosystem scenario modelling into
reef restoration decision support. This platform allows for different intervention
priorities to be dynamically targetted in decision making, while modelling the evolving
outcomes as a result of interventions. The capability of ADRIA-CoralBlox could be further
expanded by incorporating the impacts of waves, cyclones, CoTS and currents into future
resilience investigations [@lin2023; @mcclanahan2005; @grimaldi2023; @grimaldi2022;
@liu2024]. Additionally, the investigation could be enhanced by incorporating ocean
acidification projections into analyses of coral carbonate budgets [@eyre2018].

>>>>

Additionally, it is currently not possible to obtain highly accurate and precise estimates
of reef depth at the GBR-scale. Data adopted for the given assessment are said to be
accurate up to 15m from Mean Sea Level [@GBRMPA2020], however there are many sources of
uncertainty in the depth estimates used for this study. Satellite-derived estimates are
subject to different bottom types (rock, sand, seagrass, etc) and how these affect
reflectance, the water penetration depth of satellite sensors, influence of local conditions
(e.g., turbidity and wave activity), among many other factors. The median depths for each
indicative reef area were extracted from bathymetry data using zonal statistics and depth
values beyond the stated 15m accuracy threshold [@GBRMPA2020] were included. The resulting
depth estimates are therefore highly uncertain and are only adopted as a useful indication
for the purpose of the presented analysis. They are not intended to be an accurate and
precise report of reef depths across the GBR.

In addition to the usual challenges of deriving data from satellite imagery, it is
acknowledged that there is high potential for misalignment between the bathymetry data and
GBRMPA reef outlines. Hammerton and Lawrey [-@hammerton], for example, identified 120 false
positives, where features identified as reefs were in actuality sand banks or rocky reefs
(rather than coral reefs). Reefs with bathymetry data of less than 5% of its indicative area
were excluded from analysis to ensure analyses were not conducted on reefs with no data
coverage (see @sfig-depth-cutoff). The existing GBRMPA reef outlines were adopted to align
with other ongoing modelling and analyses conducted as part of the Reef Restoration and
Adaptation Program \[RRAP; <https://gbrrestoration.org>\] **\[TODO: PROPER CITATION OF
RRAP\]**.

The modelling conducted with ADRIA-CoralBlox itself is subject to many uncertainties. For
one, a formal mechanistic and spatial validation of the model represents an area for future
refinement, and is the focus of a future publication. Indicative results are included in the
supplementary **\[TI: For Pedro to include! :`D`\].** The consistency of projected patterns
with known coral-climate relationships and reef characteristics provides confidence in the
model's ability to meaningfully capture ecological signals at the GBR-scale. The mechanistic
basis of the model, combined with qualitative assessment of model behavior, supports the
robustness of our findings for understanding broad-scale coral cover responses to climatic
drivers. The aim here, however, was to assess the relative influence of reef characteristics
on structural resilience under future conditions, rather than provide spatially explicit and
highly accurate and precise predictions of reef outcomes. Using this approach we suggest
that reef depths is an important factor for reef resilience under thermal stress.

Improvements in remote sensing approaches for broad scale bathymetry measurement, and
understanding of how depth impacts experienced thermal stress on reefs will further enhance
investigations of depth as a critical factor for reef resilience. Additional analyses could
build on the findings of this study by assessing the impacts of prioritising reef depth in
reef restoration decision making. Future analyses could incorporate further environmental
and ecological stressors such as wave activity, ocean acidification, and Crown-of-Thorn
Starfish [@lin2023; @mcclanahan2005; @grimaldi2023; @grimaldi2022; @liu2024; @eyre2018].
